---
title: "portfolio 5"
output: html_document
---
Read in data
```{r}
all_data=read.csv("sockult_data.csv")

#exclude one participant with "other" as gender to run the models with it
all_data= all_data[-186,]

#only happy and neutral conditions
emotion_data= na.omit(all_data)


```

Models for willingness and eyes
```{r}
library(rethinking)

#using id as a random intercept to allow individual variation

#just alpha
#see if there was a general preference for donating over not donating
m_wiley_alpha <- map2stan(
  
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a,
    a ~ dnorm(0,10)
  ) ,
  data=all_data,
  chains=2, cores=2,iter = 5000)


plot(m_wiley_alpha)

#adding eyes - testing the hypothesis
#weakly informative prior
m_wiley_e <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*eyes_n,
    a ~ dnorm(0,10),
    be ~ dnorm(0,10)
  ) ,
  data=all_data,
  chains=2 , cores=2, iter = 5000)

all_data$id= as.numeric(all_data$id)

#adding random effect
m_wiley_er <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a[id] + be*eyes_n,
    a[id] ~ dnorm(0,10),
    be ~ dnorm(0,10)
  ) ,
  data=all_data,
  chains=2 , cores=2, iter = 5000)

#conservative prior
m_wiley_epr <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*eyes_n,
    a ~ dnorm(0,10),
    be ~ dnorm(0,1)
  ) ,
  data=all_data,
  chains=2 , cores=2, iter = 5000)

compare(m_wiley_alpha, m_wiley_e, m_wiley_er, m_wiley_epr)
compare(m_wiley_alpha, m_wiley_e, m_wiley_epr)
#adding eyes helps

results=precis(m_wiley_er, depth = 2)
get_b=results@output
get_b[206:207,]


logistic(5.21) #0.99

#relative effect
#proportional change in odds
exp(5.21) #183 -> 83%


#absolute effect
logistic(-11.34) # 1.188763e-05
logistic( -11.34 + 5.21) #0.002171854
#absolute change of less than 0.02%
```

WAIC result with random intercept:
               WAIC pWAIC dWAIC weight    SE  dSE
m_wiley_er     94.8  31.9   0.0      1  2.74   NA
m_wiley_alpha 254.1   1.0 159.3      0 11.25 9.24
m_wiley_epr   255.8   1.9 161.0      0 11.34 9.36
m_wiley_e     256.1   2.0 161.3      0 11.34 9.36

WAIC result without random intercept:
               WAIC pWAIC dWAIC weight    SE  dSE
m_wiley_alpha 254.1   1.0   0.0   0.56 11.25   NA
m_wiley_epr   255.8   1.9   1.8   0.23 11.34 0.26
m_wiley_e     256.1   2.0   2.0   0.20 11.34 0.29

using eyes only makes the model better if individual variation is allowed

Results:
       Mean   StdDev lower 0.89 upper 0.89    n_eff      Rhat
be 5.214302 1.102471   3.357982   6.852248 2874.142 0.9998596


```{r}
#using id as a random intercept to allow individual variation

#just alpha
#see if there was a general preference for donating over not donating
m_wilem_alpha <- map2stan(
  
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a,
    a ~ dnorm(0,10)
  ) ,
  data=emotion_data,
  chains=2, cores=2,iter = 5000)


plot(m_wilem_alpha)

#adding eyes - testing the hypothesis
#weakly informative prior
m_wilem_e <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*emotion_n,
    a ~ dnorm(0,10),
    be ~ dnorm(0,10)
  ) ,
  data=emotion_data,
  chains=2 , cores=2, iter = 5000)

emotion_data$id= as.numeric(emotion_data$id)

#adding random effect
m_wilem_er <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a[id] + be*emotion_n,
    a[id] ~ dnorm(0,10),
    be ~ dnorm(0,10)
  ) ,
  data=emotion_data,
  chains=2 , cores=2, iter = 5000)

#conservative prior
m_wilem_epr <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*emotion_n,
    a ~ dnorm(0,10),
    be ~ dnorm(0,1)
  ) ,
  data=emotion_data,
  chains=2 , cores=2, iter = 5000)


compare(m_wilem_alpha, m_wilem_e, m_wilem_er, m_wilem_epr )
compare(m_wilem_alpha, m_wilem_e, m_wilem_epr)
#adding eyes helps only with random intercept

results=precis(m_wilem_er, depth = 2)
get_b=results@output
get_b[142:143,]


logistic(5.82) #0.99

#relative effect
#proportional change in odds
exp(5.82) #336 -> ????%


#absolute effect
logistic(7.86) # 0.9996143
logistic( 7.86 + 5.82) #0.9999989
#absolute change of less than 0.004%
```
WAIC result with random intercept:
               WAIC pWAIC dWAIC weight   SE  dSE
m_wilem_er     64.2  21.4   0.0      1 2.19   NA
m_wilem_alpha 176.2   1.0 112.0      0 9.33 8.02
m_wilem_epr   177.7   1.9 113.4      0 9.39 7.88
m_wilem_e     177.9   2.1 113.7      0 9.50 7.96

WAIC result without random intercept:
               WAIC pWAIC dWAIC weight   SE  dSE
m_wilem_alpha 176.2   1.0   0.0   0.52 9.33   NA
m_wilem_epr   177.7   1.9   1.5   0.25 9.39 1.15
m_wilem_e     177.9   2.1   1.7   0.22 9.50 1.29

using emotions only makes the model better if individual variation is allowed

Results:
           Mean   StdDev lower 0.89 upper 0.89    n_eff      Rhat
a[142] 7.861709 6.153088 -0.8187022  17.327108 3458.016 1.0002566
be     5.821657 1.527852  3.5841335   8.477511 3154.127 0.9997625


From the book - I can't make it work
```{r}


part <- 1 
d.pred <- list( willingness_n = rep( 0 , 2 ), # empty outcome 
                eyes_n = c(0,1), # no/yes
                id = rep(part,2) ) 

link.m10.4 <- link( m_wiley_er , data=d.pred ) 
pred.p <- apply( link.m10.4 , 2 , mean ) 
pred.p.PI <- apply( link.m10.4 , 2 , PI )

plot( 0 , 0 , type="n" , xlab="eyes" , ylab="proportion willingness" ,
      ylim=c(0,1) , xaxt="n" , xlim=c(1,2) , yaxp=c(0,1,2) ) 
axis( 1 , at=1:2 , labels=c("0/0","1/1") ) 
mtext( paste( "id" , part ) )

p <- by( all_data$willingness_n , list(all_data$eyes_n,all_data$id) , mean ) 
#there are NAs!

lines( 1:2 , as.vector(p[,,part]) , col=rangi2 , lwd=2 ) #not working
#Error in p[, , part] : incorrect number of dimensions

lines( 1:2 , pred.p )
shade( pred.p.PI , 1:2 )
```

These don't seem to make sense
```{r}
prediction <- sim( m_wiley_alpha , data=all_data, n= 1000)

bayesplot::ppc_dens_overlay(all_data$willingness_n, prediction[1:20, ])

bayesplot::ppc_ecdf_overlay(all_data$willingness_n, prediction[1:20, ])

```

