---
title: "portfolio 5"
output: html_document
---
Read in data
```{r}
all_data=read.csv("sockult_data.csv")

#only happy and neutral conditions
emotion_data= na.omit(all_data)

```

#models for willingness
Models for willingness and eyes
```{r} 
library(rethinking)

#just alpha
#see if there was a general preference for donating over not donating
m_wiley_alpha <- map2stan(
  
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a,
    a ~ dnorm(0,10)
  ) ,
  data=all_data,
  chains=2, cores=2,iter = 5000)


#plot(m_wiley_alpha)

#adding eyes - testing the hypothesis
#weakly regularizing prior
m_wiley_e <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*eyes_n,
    a ~ dnorm(0,10),
    be ~ dnorm(0,10)
  ) ,
  data=all_data,
  chains=2 , cores=2, iter = 5000)

all_data$id= as.numeric(all_data$id)

#adding random intercept
#allows for individual variation
m_wiley_er <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a[id] + be*eyes_n,
    a[id] ~ dnorm(0,10),
    be ~ dnorm(0,10)
  ) ,
  data=all_data,
  chains=2 , cores=2, iter = 5000)

#conservative prior
m_wiley_epr <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*eyes_n,
    a ~ dnorm(0,10),
    be ~ dnorm(0,1)
  ) ,
  data=all_data,
  chains=2 , cores=2, iter = 5000)

#informative prior
#expecting positive effect
m_wiley_eipr <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*eyes_n,
    a ~ dnorm(0,10),
    be ~ dnorm(5,10)
  ) ,
  data=all_data,
  chains=2 , cores=2, iter = 5000)

compare(m_wiley_alpha, m_wiley_e, m_wiley_er, m_wiley_epr,m_wiley_eipr)
compare(m_wiley_alpha, m_wiley_e, m_wiley_epr, m_wiley_eipr)
#using eyes only makes the model better if individual variation is allowed

#to get only b leave out depth
results=precis(m_wiley_er, depth = 2)
get_b=results@output
x=get_b[c(20, 33, 64, 96, 103, 132, 150, 190, 200, 203, 209),] #just random numbers


logistic(5.21) #0.99

#relative effect
#proportional change in odds
exp(5.21) #183 -> 83%


#absolute effect
logistic(-11.34) # 1.188763e-05
logistic( -11.34 + 5.21) #0.002171854
#absolute change of less than 0.02%

#absolute effect
logistic(4.8) # 0.9918374
logistic( 4.8 + 5.21) #0.9999551
#absolute change of less than 0.08%

#plot(precis(m_wiley_er, depth = 2))
```

WAIC result with random intercept:
               WAIC pWAIC dWAIC weight    SE  dSE
m_wiley_er     93.9  31.4   0.0      1  2.60   NA
m_wiley_alpha 257.1   1.0 163.1      0 11.21 9.38
m_wiley_epr   259.1   2.0 165.1      0 11.30 9.48
m_wiley_eipr  259.1   2.0 165.1      0 11.23 9.40
m_wiley_e     259.4   2.1 165.4      0 11.36 9.54

WAIC result without random intercept:
               WAIC pWAIC dWAIC weight    SE  dSE
m_wiley_alpha 257.1   1.0   0.0   0.49 11.21   NA
m_wiley_epr   259.1   2.0   2.0   0.18 11.30 0.16
m_wiley_eipr  259.1   2.0   2.0   0.18 11.23 0.11
m_wiley_e     259.4   2.1   2.3   0.15 11.36 0.21

using eyes only makes the model better if individual variation is allowed
using different priors doesn't change much

Results:
       Mean   StdDev lower 0.89 upper 0.89    n_eff      Rhat
be       5.196442 1.109334   3.494915   7.012474 2395.498 1.0023773

Models for willingness and emotion
```{r}

#just alpha
#see if there was a general preference for donating over not donating
m_wilem_alpha <- map2stan(
  
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a,
    a ~ dnorm(0,10)
  ) ,
  data=emotion_data,
  chains=2, cores=2,iter = 5000)


#plot(m_wilem_alpha)

#adding eyes - testing the hypothesis
#weakly informative prior
m_wilem_e <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*emotion_n,
    a ~ dnorm(0,10),
    be ~ dnorm(0,10)
  ) ,
  data=emotion_data,
  chains=2 , cores=2, iter = 5000)

emotion_data$id= as.numeric(emotion_data$id)

#adding random effect
m_wilem_er <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a[id] + be*emotion_n,
    a[id] ~ dnorm(0,10),
    be ~ dnorm(0,10)
  ) ,
  data=emotion_data,
  chains=2 , cores=2, iter = 5000)

#conservative prior
m_wilem_epr <- map2stan(
  alist(
    willingness_n ~ dbinom( 1 , p ) ,
    logit(p) <- a + be*emotion_n,
    a ~ dnorm(0,10),
    be ~ dnorm(0,1)
  ) ,
  data=emotion_data,
  chains=2 , cores=2, iter = 5000)


compare(m_wilem_alpha, m_wilem_e, m_wilem_er, m_wilem_epr )
compare(m_wilem_alpha, m_wilem_e, m_wilem_epr)
#adding eyes helps only with random intercept

results=precis(m_wilem_er, depth = 2)
get_b=results@output
b=get_b[c(12, 23, 57, 60, 64, 73, 75, 99, 112, 124, 140, 143),]


logistic(5.84) #0.99

#relative effect
#proportional change in odds
exp(5.84) #343 -> ????%


#absolute effect
logistic(-7.89) # 0.0003743294
logistic( -7.89 + 5.84) #0.1140524
#absolute change of less than 11.4%

logistic(7.94) # 0.9996439
logistic( 7.94 + 5.84) #0.999999
#absolute change of less than 0.004%

```
WAIC result with random intercept:
               WAIC pWAIC dWAIC weight   SE  dSE
m_wilem_er     64.2  21.4   0.0      1 2.19   NA
m_wilem_alpha 176.2   1.0 112.0      0 9.33 8.02
m_wilem_epr   177.7   1.9 113.4      0 9.39 7.88
m_wilem_e     177.9   2.1 113.7      0 9.50 7.96

WAIC result without random intercept:
               WAIC pWAIC dWAIC weight   SE  dSE
m_wilem_alpha 176.2   1.0   0.0   0.52 9.33   NA
m_wilem_epr   177.7   1.9   1.5   0.25 9.39 1.15
m_wilem_e     177.9   2.1   1.7   0.22 9.50 1.29

using emotions only makes the model better if individual variation is allowed

Results:
           Mean   StdDev lower 0.89 upper 0.89    n_eff      Rhat
be       5.842639 1.504150   3.4636500   8.211291 2743.772 0.9997234


#plots for willingness

```{r prior check} 


prior_check= function(model,p, s) {
  #plotting the priors against the posteriors
  post = extract.samples( model , n=1e4 )
  dist= data.frame(poste= p, prior= rnorm(1e4, 0, s)) #distribution of alpha/beta (posterior and prior)
  
  ggplot(data= dist)+ 
    geom_line(aes(poste), stat= "density") +
    geom_line(aes(prior), stat= "density", colour= "red")+
    labs(x="data", y="density")
  
}

#prior is red

prior_check(m_wilem_e,post$a, 10)
prior_check(m_wilem_e,post$be, 10)
prior_check(m_wilem_epr,post$be, 1)
prior_check(m_wilem_epr,post$a, 1)

prior_check(m_wiley_e,post$a, 10)
prior_check(m_wiley_e, post$be, 10)
prior_check(m_wiley_epr,post$a, 1)
prior_check(m_wiley_epr, post$be, 1)
#seems fine?
```


```{r coeftab} 
# coeftab(m_wiley_alpha, m_wiley_e, m_wiley_er, m_wiley_epr, m_wiley_eipr )

plot(coeftab(m_wiley_alpha, m_wiley_e, m_wiley_epr, m_wiley_eipr))

# coeftab(m_wilem_alpha, m_wilem_e, m_wilem_er, m_wilem_epr)

plot(coeftab(m_wilem_alpha, m_wilem_e, m_wilem_epr))

# plot(coeftab(m_wilem_er)) #not good idea

```



```{r eyes}  

#get out betas and their 89%PI from the 3 models

#for models without random intercept
betas=function(x) {
  e=precis(x)#save output of precis
  s=as.character(x@formula[2]) #save equation for p
  d=as.character(x@formula[4]) #save prior for beta
  f=paste(s, ",", d) #put together to use as name
  
  r=e@output$Mean
  t=e@output$`lower 0.89`
  o=e@output$`upper 0.89`
  u=data.frame(b=r, lower=t, upper=o, name=f)
  u=u[2,] #saves alpha too, so delete first row (alpha)
  return(u)
}

#for the random intercept model
betas_r=function(x) {
  e=precis(x)
  s=as.character(x@formula[2])
  d=as.character(x@formula[4])
  f=paste(s, ",", d)
  
  r=e@output$Mean
  t=e@output$`lower 0.89`
  o=e@output$`upper 0.89`
  u=data.frame(b=r, lower=t, upper=o, name=f)
  return(u)
}

wiley_e_res=betas(m_wiley_e)
wiley_epr_res=betas(m_wiley_epr)
wiley_er_res= betas_r(m_wiley_er)

wiley_res=rbind(wiley_e_res, wiley_epr_res, wiley_er_res)

library(ggplot2)

ggplot(wiley_res) +
  geom_pointrange(aes(x=name, y=b, ymin=lower, ymax=upper), colour= "darkblue") +
  labs(x= "model", y="beta") 

```

```{r emotion} 

#get out betas and their 89%PI from the 3 models

#use functions made for eyes
wilem_e_res=betas(m_wilem_e)
wilem_epr_res=betas(m_wilem_epr)
wilem_er_res= betas_r(m_wilem_er)

wilem_res=rbind(wilem_e_res, wilem_epr_res, wilem_er_res)



ggplot(wilem_res) +
  geom_pointrange(aes(x=name, y=b, ymin=lower, ymax=upper), colour= "darkblue") +
  labs( x= "model", y="beta") +
  theme_minimal()

```

#models for amount
Models for amount and eyes
```{r} 
#from book
 # discrete proportion of each response value 
pr_k <- table( all_data$amount_n ) / nrow(all_data)
# cumsum converts to cumulative proportions 
cum_pr_k <- cumsum( pr_k )
# # plot 
# plot( 1:6 , cum_pr_k , type="b" , xlab="amount" , ylab="cumulative proportion" , ylim=c(0,1) )

#values needed for the models
logit <- function(x) log(x/(1-x)) 
(lco <- logit( cum_pr_k))
#-1.8088539 -0.2933478  0.5787368  1.1913940  1.2465324

```

```{r}
#just alpha
m_amey_alpha <-
  map2stan(
    alist(
      amount_n ~ dordlogit(phi , cutpoints),
      phi <- 0,
      cutpoints ~ dnorm(0, 10)
    ) ,
    data = list(amount_n = all_data$amount_n),
    start = list(cutpoints = c(-1.8, -0.3, 0.6, 1.19, 1.24)) ,
    chains = 2,cores = 2, iter= 5000 )

#adding eyes
m_amey_e <-
  map2stan(
    alist(
      amount_n ~ dordlogit(phi , cutpoints),
      phi <- be* eyes_n,
      be ~ dnorm(0,10),
      cutpoints ~ dnorm(0, 10)
      
    ) ,
    data = list(amount_n = all_data$amount_n, eyes_n= all_data$eyes_n),
    start = list(cutpoints = c(-1.8, -0.3, 0.6, 1.19, 1.24)) ,
    chains = 2,cores = 2, iter= 5000 )


#plot(m_amey_e)


#conservative prior
m_amey_e_pr <-
  map2stan(
    alist(
      amount_n ~ dordlogit(phi , cutpoints),
      phi <- be* eyes_n,
      be ~ dnorm(0,1),
      cutpoints ~ dnorm(0, 10)
      
    ) ,
    data = list(amount_n = all_data$amount_n, eyes_n= all_data$eyes_n),
    start = list(cutpoints = c(-1.8, -0.3, 0.6, 1.19, 1.24)) ,
    chains = 2,cores = 2, iter= 5000  )


compare(m_amey_alpha, m_amey_e, m_amey_e_pr)
#conservative prior is best

precis(m_amey_e_pr,depth=2)
plot(precis(m_amey_e_pr,depth=2))

```

               WAIC pWAIC dWAIC weight    SE   dSE
m_amey_e_pr  680.0   5.8   0.0   0.55 12.05    NA
m_amey_e     680.4   6.0   0.4   0.45 12.08  0.14
m_amey_alpha 864.4  15.1 184.4   0.00 30.59 33.29

             Mean StdDev lower 0.89 upper 0.89 n_eff Rhat
cutpoints[1] -1.93   0.27      -2.37      -1.51  3760    1
cutpoints[2] -0.40   0.22      -0.75      -0.04  4697    1
cutpoints[3]  0.47   0.23       0.09       0.81  5000    1
cutpoints[4]  1.12   0.24       0.73       1.50  5000    1
cutpoints[5]  1.20   0.24       0.78       1.56  5000    1
be           -0.12   0.26      -0.54       0.28  4461    1

Models for amount and emotion
```{r}
#from book
 # discrete proportion of each response value 
pr_k <- table( emotion_data$amount_n ) / nrow(emotion_data)
# cumsum converts to cumulative proportions 
cum_pr_k <- cumsum( pr_k )
# # plot 
# plot( 1:6 , cum_pr_k , type="b" , xlab="amount" , ylab="cumulative proportion" , ylim=c(0,1) )

#values needed for the models
logit <- function(x) log(x/(1-x)) 
(lco <- logit( cum_pr_k))
#-1.8082888 -0.2263131  0.5485660  1.1948403  1.2755430  
```

```{r}
#just alpha
m_amem_alpha <-
  map2stan(
    alist(
      amount_n ~ dordlogit(phi , cutpoints),
      phi <- 0,
      cutpoints ~ dnorm(0, 10)
    ) ,
    data = list(amount_n = emotion_data$amount_n),
    start = list(cutpoints = c(-1.8, -0.2, 0.5, 1.19, 1.27)) ,
    chains = 2,cores = 2, iter= 5000 )

#adding eyes
m_amem_e <-
  map2stan(
    alist(
      amount_n ~ dordlogit(phi , cutpoints),
      phi <- be* eyes_n,
      be ~ dnorm(0,10),
      cutpoints ~ dnorm(0, 10)
      
    ) ,
    data = list(amount_n = emotion_data$amount_n, eyes_n= emotion_data$eyes_n),
    start = list(cutpoints = c(-1.8, -0.2, 0.5, 1.19, 1.27)) ,
    chains = 2,cores = 2, iter= 5000 )

#plot(m_amem_e)


#conservative prior
m_amem_e_pr <-
  map2stan(
    alist(
      amount_n ~ dordlogit(phi , cutpoints),
      phi <- be* eyes_n,
      be ~ dnorm(0,1),
      cutpoints ~ dnorm(0, 10)
      
    ) ,
    data = list(amount_n = emotion_data$amount_n, eyes_n= emotion_data$eyes_n),
    start = list(cutpoints = c(-1.8, -0.2, 0.5, 1.19, 1.27)) ,
    chains = 2,cores = 2, iter= 5000  )


compare(m_amem_alpha, m_amem_e, m_amem_e_pr)

precis(m_amem_e_pr,depth=2)
plot(precis(m_amem_e_pr,depth=2))

```

              WAIC pWAIC dWAIC weight    SE   dSE
m_amem_e_pr  468.6   4.8   0.0   0.52 10.55    NA
m_amem_e     468.7   4.9   0.1   0.48 10.63  0.10
m_amem_alpha 584.0  11.4 115.4   0.00 22.62 25.45


              Mean StdDev lower 0.89 upper 0.89 n_eff Rhat
cutpoints[1] -1.88   1.00      -3.53      -0.32  2052    1
cutpoints[2] -0.28   0.99      -1.82       1.31  2102    1
cutpoints[3]  0.52   0.99      -1.00       2.14  2139    1
cutpoints[4]  1.18   1.00      -0.42       2.74  2150    1
cutpoints[5]  1.30   1.00      -0.30       2.89  2149    1
be           -0.04   0.98      -1.50       1.61  2128    1


#plots for amount
```{r coeftab}

coeftab(m_amey_alpha, m_amey_e, m_amey_e_pr)

plot(coeftab(m_amey_alpha, m_amey_e, m_amey_e_pr))


coeftab(m_amem_alpha, m_amem_e, m_amem_e_pr)

plot(coeftab(m_amem_alpha, m_amem_e, m_amem_e_pr))



```





